---
title: "HW_1"
author: "Stanislav Sochynskyi"
date: "March 2, 2019"
output:
  html_document:
    df_print: paged
  pdf_document: default
---
File import
```{r}
df<-read.csv("inputfile1.csv",sep=" ")
df
```

Task #1. Identify the data type of each column. Write the command and its corresponding output.
```{r}
print(class(df$time))
print(class(df$account_a))
print(class(df$account_b))
print(class(df$account_c))
```
Task #2. For each columnin dataset:
a) Find unique values in case column is oftype factor;
```{r}
unique(df$time)
```
b) Find ranges(minimum and maximum)if column belongs to numeric type:
For column "account_a".
```{r}
min(df$account_a) 
max(df$account_a) 
range(df$account_a) 
```
For column "account_b".
```{r}
min(df$account_b) 
max(df$account_b) 
range(df$account_b) 
```
For column "account_c".
```{r}
min(df$account_c) 
max(df$account_c) 
range(df$account_c) 
```

Task #3. Build 3 scatter plots in order to find correlation among threebank accounts. Is there any relation between any 2 bank accounts?

We need to load "ggplot2" library for this task.
```{r}
library("ggplot2")
```

```{r}
ggplot(df)+
  geom_point(aes(x=df$account_a,y=df$account_b))
cor(df$account_a, df$account_b)
```
Negative correlation between "account_a"" and "account_b"" is observed in the grap. Function "cor" showed "-0.9671627" which proves our observation from the scatter plot.

```{r}
ggplot(df)+
  geom_point(aes(x=df$account_a,y=df$account_c))
cor(df$account_a, df$account_c)
```
There's no correlation between "account_a"" and "account_c". Function "cor" showed very low negative number = -0.02937619. 

```{r}
ggplot(df)+
  geom_point(aes(x=df$account_c,y=df$account_b))
cor(df$account_c, df$account_b)
```
There's no correlation between "account_a"" and "account_c". Function "cor" showed very low positive number = 0.005483397. 

Task #4. 
Before selecting records
Select records for 2016 year only. Before selecting records for 2016 only, we need to . We can either use library "lubridate" or function as.Date. Let's stop with a secind option. To verify that let's use the function "class" to verify that column "time" is not
```{r}
df_copy<-df
df_copy$time<-as.Date(df$time, format="%Y-%m-%d")
class(df_copy$time)
```
Now, let's select the 2016 records and sort them in descending order by time. For that we will need function filter from library "dplyr.
```{r}
library(dplyr)
library(reshape2)
#*  select(): choosing a subset of columns
#*  filter(): choosing a subset of rows
accounts_2016<- df_copy %>%
  filter(time >= "2016-01-01" & time <="2016-12-01") %>%
  arrange(desc(time))%>%
  melt(id="time") #transform to long format based in time
```
Build a single plot which will show how balance changed for all 3 accounts during that year. Use line plot for this task. (Each account should have different color).
```{r}
ggplot(accounts_2016, aes(x=time, y=value))+
            geom_line(aes(color=variable))
```

Task #5. Group data by the year, for each year and account calculate average balance. 
To do it we need to use function format() extract year from the time column.
```{r}
#class(accounts$time)
df_copy$time<-format(df_copy$time,"%Y")
grouped_data = group_by(df_copy, time)
summary<-summarise(grouped_data, mean(account_a), mean(account_b), mean(account_c))
summary
```

```{r}
summary<-melt(summary,id="time")
```

Build a heat map to see how average balance changed on each account during the years. What can you say based on the plotted heat map? 
```{r}
ggplot(summary, aes(x=time, y=variable))+
  geom_tile(aes(fill=value))
```

Write at least 3 observations from the heatmap.
1. The average balance of account c hasn't changed significantly for the past 10 years.
2. The average balance of account b increased over the 10 years.
3. The average balance of account a decreased for the past 10 years.
4. Heatmap also shows a negative correlation between account a and account b.

