---
title: "Untitled"
author: "Stanislav Sochynskyi, Katsiaryna Lashkevich"
date: "April 30, 2019"
output: html_document
---

Here we are! The last homework... 
And what a beutiful end!

>>Task 2.1  Calculate statistical information

As always let's first read the files and add required libraries for analysis.
```{r}
library("igraph")
library("ggplot2")

edges<- read.csv("Hi-tech-Edges.csv",sep=",")
nodes<-read.csv("Hi-tech-Nodes.csv",sep=",")
```

Let's start to build graph rightaway. However, first let's check edges for uniqueness and sum up in case we have duplicates.

```{r}
#check for uniqueness
cat("Amount of rows in edges file: ", nrow(edges), "\n")
cat("Amount of unique edges: ", nrow(unique(edges[,c("from", "to")])), "\n")
```
The number of rows in edges is the same as number of unique words in the document thus we can conclude there is no duplicates.

Let's first calculate statistical information about our net such as _density_, _clustering coefficient_, _Reciprocity_, _average path length_, _diameter (by considering weights)_.

```{r}
net <- graph_from_data_frame(d=edges, vertices=nodes, directed = T)
#density
edge_density(net)
#clutering coefficient
transitivity(net, type="global") #average of local

transitivity(net, type="local")
#reciprocity of the network
reciprocity(net)

#average path length
mean_distance(net, directed = TRUE, unconnected = TRUE)

#diameter
diameter(net, directed=T)
```
1. Density ratio is only *~0.1*. This means that the network is not tightly connected.
2. The ratio of the number of loops of length three and the number of paths of length two *~0.35*
3. As we have directly connected network we can calculate *reciprocity* of the network. The calue is *~0.67* which means that more than a half of nodes have edges in both directions ovar the total number of relations in the network. Basically, it means that many people have connections with each other.
4. This parameter shows that the distance between all pairs on average is 2.63 
5. Length of the shortest path between two nodes (diameter) is 102.

Let's now build the graph from the data!

>>Task 2.2  Plot the graph with following rules.

Assigning shapes to the Gender variable:
```{r}
shape <- c("square", "circle")
V(net)$shape=shape[as.numeric(as.factor(V(net)$Gender))]
V(net)$shape
```
Assigning colors to the Department variable:
```{r}
#unique(V(net)$Department) - 3 unique department: Marketing, Management, Development

library(RColorBrewer)
coul = brewer.pal(3, "Set1") 
V(net)$color=coul[as.numeric(as.factor(V(net)$Department))]
#V(net)$color
```
Assigning size to the edges depending on their weight:
```{r}
E(net)$width <- E(net)$weight/8
```
Let's find the hub size of the nodes and assign the size:
```{r}
hs <- hub_score(net, weights=NA)$vector
```
Let's find the diameter and colorize its edges:
```{r}
diam <- get_diameter(net, directed=T)
vcol <- V(net)$color
vcol[diam] <- "yellow"
ecol <- rep("gray80", ecount(net))
ecol[E(net, path=diam)] <- "yellow"
```
Now, let's plot the net!
```{r}
plot(net, 
     edge.arrow.size = .2, 
     vertex.label.dist = 2, 
     vertex.label.cex = 0.8,
     vertex.size = hs*40,
     vertex.label = V(net)$Name,
     vertex.color= V(net)$color,
     vertex.label.color = "black",
     edge.color=ecol,
     layout= layout.grid(net,width = 6))
```

>> Task 2.3 

In order to find the number of the communities we use the Walktrap algorithm:
```{r}
wtc <- walktrap.community(as.undirected(net))
wtc
```

Plotting the communities:
```{r}
plot(wtc, net, 
     edge.arrow.size = .2, 
     vertex.label.dist = 2, 
     vertex.label.cex = 0.8,
     vertex.size = hs*40,
     vertex.label = V(net)$Name,
     vertex.color= V(net)$color,
     vertex.label.color = "black",
     edge.color=ecol,
     main = "Hubs",
     layout= layout.grid(net,width = 6))
```

It is seen from the output that there are 11 communities. 
Let's take a look at some samples. For instance, nodes 1 (Abe) and 6 (Fred) have connections only with each other that's why they appeared in one group - Group 2. Group 7 includes nodes 3 (Carl) and 8 (Helen), which also have connections only with each other. Nodes 3 (Carl) and 8 (Helen) do not have connections with nodes 1 (Abe) and 6 (Fred), that means it is right they appear in different groups.